---
title: "SMTrackR"
author: "Aashna Bansal, Satyanarayan Rao"
date: "2025-06-12"
vignette: >
      %\VignetteIndexEntry{SMTrackR}
      %\VignetteEngine{knitr::rmarkdown}
      \usepackage[utf8]{inputenc}
output: 
  BiocStyle::html_document:
    toc: true
    toc_depth: 2

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Single-molecule assays like NOMe-seq, and dSMF are superior to DNase-seq and ATAC-seq as they do not nibble down DNA. Thus, they enable quantification of all three i.e., protein-free, Transcription Factor-bound and histone-complex-bound states. But a user-friendly tool to visualize and quantify such states is lacking. SMTrackR, a Bioconductor package visualizes protein-DNA binding states on individual sequenced DNA molecules. SMTrackR queries the single-molecule footprint database we built and hosted at Galaxy Server. It comprises of BigBed files generated from NOMe-seq and dSMF datasets. SMTrackR exploits UCSC REST API to query a BigBed file and plot footprint heatmap categorized in different binding states as well as report their occupancies.  

## Installation
```{r, echo = T}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("SMTrackR")
# rmarkdown::html_vignette
```
## Loading the package
```{r Load package, echo=TRUE}
library(SMTrackR)
```
## Plot Single Molecules from dSMF Data

```{r Plot dSMF data, fig.width=4.5, fig.height=6, echo=TRUE}
SMTrackR::plotFootprints()
```

## Plot Single Molecules from SMF Data

`SMTrackR::plotFootprints` will generate the heatmap `pdf` file stored in `plots/<label>.plot.pdf`. 
A `Gviz` compatible code can also be generated by running the `generateGvizCodeforSMF` function. See the code chunk below. 

```{r plot psuedo-bulk scNOMe-seq data, fig.width=4.5, fig.height=6, echo=TRUE}
SMTrackR::plotFootprints(organism = "mmusculus", model = "8cell", condition = "WT", 
                         genome_assembly = "mm10", type = "SMF", chromosome = "chr5", 
                         start = 113847750,  stop  = 113847780, tr = "8cell", 
                         label = "check", fp_cap = 50, remove_dup = T)
SMTrackR::generateGvizCodeforSMF(organism = "mmusculus", model = "8cell", condition = "WT",
                                 genome_assembly = "mm10", type = "SMF", chromosome = "chr5", 
                                 start = 113847750,  stop  = 113847780, tr = "8cell", 
                                 label = "check", fp_cap = 50, remove_dup = T,
                                 gviz_left = 1000, gviz_right = 1000)


```

## Plot Single Molecules from Nanopore Data

```{r, fig.width=10, fig.height=5, echo=TRUE}
SMTrackR::plotMethylationCallsNanopore() 
```

## Listing Available Track
```{r List available tracks, echo = TRUE}

SMTrackR::listTracks()
```



## Session Info
```{r, echo = TRUE}
sessionInfo()
```
