---
title: "SMTrackR"
author: "Aashna Bansal, Satyanarayan Rao"
date: "2025-06-12"
output: 
  BiocStyle::html_document:
    toc: true
    toc_depth: 2

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Single-molecule assays like NOMe-seq, and dSMF are superior to DNase-seq and ATAC-seq as they do not nibble down DNA. Thus, they enable quantification of all three i.e., protein-free, Transcription Factor-bound and histone-complex-bound states. But a user-friendly tool to visualize and quantify such states is lacking. SMTrackR, a Bioconductor package visualizes protein-DNA binding states on individual sequenced DNA molecules. SMTrackR queries the single-molecule footprint database we built and hosted at Galaxy Server. It comprises of BigBed files generated from NOMe-seq and dSMF datasets. SMTrackR exploits UCSC REST API to query a BigBed file and plot footprint heatmap categorized in different binding states as well as report their occupancies.  
## Installation
```{r, echo = T}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("SMTrackR")
```
## Loading the package
```{r, echo = T}
library(TestSMTR)
```
## Listing Available Track
```{r plotting, echo = TRUE}

TestSMTR::listTracks()
```

## Plot Single Molecules
```{r, fig.width=4.5, fig.height=6, echo=TRUE}
TestSMTR::plotFootprints()
```

